Time Series

### 2023-11-17 LPR/index.qmd
Updates

```{r shop-data}
library(tibble)
library(dplyr)

cutting_surfaces <- tribble(
  ~type, ~length, ~surfaces, ~tools,
  "Pliers", 1.5, 2, 10,
  "Needlenose pliers", 1, 4, 3,
  "Side cutters", 1, 4, 1,
  "Side cutters", 2, 4, 2,
  "Scissors", 8.89, 2, 14,
  "Tin snips", 3.81, 2, 4,
  "Lineman pliers", 2, 4, 2,
  "Lineman pliers", 1, 4, 1,
  "Wire strippers", 1, 2, 2,
  "Long handled cutting pliers", 2, 4, 1,
  "Long handled cutting pliers", 3, 4, 1,
  "Utility knife blades", 7.62, 2, 50
) %>%
  mutate(total_cutting_length = length*surfaces*tools)

total_cutting_surface = sum(cutting_surfaces$total_cutting_length)
```

@@

A scan of a cut made by the blade at $0.625\mu m$ resolution would produce approximately $15/0.0000625 = 240000$ observations; a scan of a 2mm wire would produce $2/0.0000625 = 32000$ observations, and computationally aligning these two data sequences using maximum cross-correlation would require $240000 - 32000 + 1 = 208001$ comparisons. Under an optical microscope, an examiner might move the wire at 0.2mm increments along the blade cut surface for a total of $15/.2 - 2/.2 + 1 = 66$ distinct and discrete comparisons. Then, we must consider that the examiner or algorithm must compare the wire to both side A and side B; we then double the number of total comparisons made to $416002$ and $132$ for the algorithm and the examiner, respectively.

@@

While 132 visual comparisons might not seem so terrible, this is the simplest of the 3 scenarios we consider; each scenario presents a progressively more complex problem.

@@

Let us consider the garage of one of the authors of this paper; @fig-shop-layout shows the garage work bench and tool storage in typical condition. While there is no reason to think the author's garage is representative of all garages, the general setup and number of tools in the garage does not seem unusual for a suburban household where residents may dabble in automotive repair, home improvement, or woodworking.

@@

```{r}
#| label: tbl-shop-surfaces
#| tbl-cap: "Shop cutting surfaces. All lengths in cm."
#| echo: false
#| message: false
knitr::kable(cutting_surfaces)
```

```{r}
#| label: tbl-shop-comparisons
#| tbl-cap: "Comparions performed by tool type, assuming digital scans at a resolution of $0.625\\mu m$ and visual comparisons performed at $0.2 mm$ intervals along the test blade impression."
#| echo: false
#| message: false
comparisons <- cutting_surfaces %>%
  mutate(digital = (length*10/0.0000625 - 2/0.0000625+1)*surfaces*tools,
         analog = (length*10/.2 - 2/.2 + 1)) %>%
  group_by(type) %>%
  summarize(digital = sum(digital), analog = sum(analog)) %>%
  rename(`Tool type` = type, Algorithm = digital, Examiner = analog)

knitr::kable(comparisons)
```

We assembled the easily available tools from the house and garage which might reasonably be used to cut wires[^2], including pliers, wire cutters and strippers, scissors, tongs, and utility knives. Estimated cutting surface length and quantity of each type of tool is provided in @tbl-shop-surfaces. The total length of all cutting surfaces of relevant tools in the author's garage is `r total_cutting_surface` cm; primary contributions to the total length are the 50 pack of utility knife blades and the 14 pairs of full-size scissors which could be located throughout the house. Of course, a thorough search of the garage and house would come up with far more cutting surfaces to compare against during an actual crime investigation.

[^2]: Reasonably is defined as we have used a tool like this to cut wires during electrical work in the past.

As in Scenario 1, the wire fragment recovered from the un-detonated explosive would be compared to blade surface test impressions generated from each cutting surface in the garage. However, unlike in Scenario 1, we have `r total_cutting_surface` cm of cutting surfaces to work with. We can apply our comparison formula on a by-surface basis to yield the digital and analog comparison estimates shown in @tbl-shop-comparisons. Overall, our algorithm would make `r sum(comparisons$Algorithm)` comparisons and an examiner doing a visual inspection would make `r sum(comparisons$Examiner)` comparisons under the assumptions we made in Scenario 1. XXX fill this in with proper formatting XXX

@@

>>>>

```{r shop-data}
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)

cutting_surfaces <- tribble(
  ~type, ~length, ~surfaces, ~tools,
  "Pliers", 1.5, 2, 10,
  "Pliers", 1, 4, 3,
  "Side cutters", 1, 4, 1,
  "Side cutters", 2, 4, 2,
  "Scissors", 8, 2, 14,
  "Tin snips", 4, 2, 4,
  "Lineman pliers", 2, 4, 2,
  "Lineman pliers", 1, 4, 1,
  "Wire strippers", 1, 2, 2,
  "Long handled cutting pliers", 2, 4, 1,
  "Long handled cutting pliers", 3, 4, 1,
  "Utility knife blades", 6.2, 2, 50
) %>%
  mutate(total_cutting_length = length*surfaces*tools)

total_cutting_surface = sum(cutting_surfaces$total_cutting_length)

dad_surfaces <- tribble(
  ~type, ~length, ~surfaces, ~tools,
  "Pliers", 1.8, 2, 1,
  "Pliers", 1.6, 2, 1,
  "Pliers", 1.1, 2, 1,
  "Wire cutters", 2, 2, 1,
  "Wire cutters", 1.5, 2, 1,
  "Wire cutters", 1, 2, 2,
  "Other cutters", 8, 2, 1,
  "Other cutters", 4, 4, 1,
  "Scissors", 7.5, 2, 5,
  "Utility knife blades", 6.2, 2, 160,
  "Razor blades", 4, 2, 14,
  "Knife", 4, 2, 1,
  "Knife", 2, 2, 1,
  "Clippers", 4, 2, 1
) %>%
  mutate(total_cutting_length = length*surfaces*tools)

dad_total_cutting_surface = sum(dad_surfaces$total_cutting_length)
```

@@

```{r}
resolution <- 0.645 # res in mum
length_blade <- 15 # length in mm
wire_diam <- 2 # diameter in mm
mm_to_mum <- 1000 # conversion

obs_blade <- round(length_blade * mm_to_mum/resolution)
obs_wire <- round(wire_diam * mm_to_mum/resolution)
alg_compare <- obs_blade - obs_wire + 1

examiner_increment <- 0.2 # increment in mm
examiner_compare <- (length_blade - wire_diam)/examiner_increment + 1
```

A scan of a cut made by the blade at $`r resolution`\mu m$ resolution will produce approximately $`r length_blade` \times `r mm_to_mum`/`r resolution` = `r format(obs_blade, big.mark = ',')`$ observations; a scan of a 2mm wire would produce $`r wire_diam` \times `r mm_to_mum`/`r resolution` = `r format(obs_wire, big.mark = ',')`$ observations, and computationally aligning these two data sequences using maximum cross-correlation would require $`r obs_blade` - `r obs_wire` + 1 = `r format(alg_compare, big.mark = ',')`$ (obviously dependent) comparisons. Under an optical microscope, an examiner might move the wire at 0.2mm increments along the blade cut surface for a total of $`r length_blade`/`r examiner_increment` - `r wire_diam`/`r examiner_increment` + 1 = `r format(examiner_compare, big.mark = ',')`$ distinct comparisons. Then, we must consider that the examiner or algorithm must compare the wire to both side A and side B; we then double the number of total comparisons made to $`r format(alg_compare*2, big.mark = ',')`$ and $`r examiner_compare*2`$ for the algorithm and the examiner, respectively.

@@

While $`r examiner_compare*2`$ visual comparisons might not seem so terrible, this is the simplest of the 3 scenarios we consider; each scenario presents a progressively more complex problem.

@@

Let us consider the garage of one of the authors of this paper; @fig-shop-layout shows the garage work bench and tool storage in typical condition. While there is no reason to think the author's garage is representative of all garages, the general setup and number of tools in the garage is not unusual for a suburban household where residents may dabble in automotive repair, home improvement, or woodworking.

@@

```{r}
#| label: tbl-shop-surfaces
#| tbl-cap: "Shop cutting surfaces. All lengths in cm."
#| echo: false
#| message: false
knitr::kable(cutting_surfaces %>% rename(Type = type, `Length (cm)` = length, `Cutting Surfaces` = surfaces, `# Tools` = tools, `Total Cutting Length (cm)` = total_cutting_length))
```

```{r}
#| label: tbl-shop-comparisons
#| tbl-cap: "Comparions performed by tool type, assuming digital scans at a resolution of $0.625\\mu m$ and visual comparisons performed at $0.2 mm$ intervals along the test blade impression."
#| echo: false
#| message: false
comparisons <- cutting_surfaces %>%
  mutate(digital = (length*10*mm_to_mum/resolution - wire_diam*mm_to_mum/resolution + 1)*surfaces*tools,
         analog = (length*10/examiner_increment - wire_diam/examiner_increment + 1)) %>%
  group_by(type) %>%
  summarize(digital = sum(round(digital)), analog = sum(round(analog))) %>%
  rename(`Tool type` = type, Algorithm = digital, Examiner = analog)

compare_tbl <- bind_rows(comparisons, tibble(`Tool type` = "Total", Algorithm = sum(comparisons$Algorithm), Examiner = sum(comparisons$Examiner)))
knitr::kable(compare_tbl, format.args = list(big.mark=','))
```

We assembled the easily available tools from the house and garage which might reasonably be used to cut wires[^2], including pliers, wire cutters and strippers, scissors, tongs, and utility knives. Estimated cutting surface length and quantity of each type of tool is provided in @tbl-shop-surfaces. The total length of all cutting surfaces of relevant tools in the author's garage is `r total_cutting_surface` cm; primary contributions to the total length are the 50 pack of utility knife blades and the 14 pairs of full-size scissors which could be located throughout the house. Of course, a thorough search of the garage and house, as one would expect from professionals investigating an actual crime, would yield far more cutting surfaces to compare against.

[^2]: Reasonably is defined as we (the authors) have used a tool like this to cut wires while doing home improvement projects in the past.

As in Scenario 1, the wire fragment recovered from the un-detonated explosive would be compared to blade surface test impressions generated from each cutting surface in the garage. However, unlike in Scenario 1, we have `r format(total_cutting_surface, big.mark=',')` cm of cutting surfaces to work with. We can apply our comparison formula on a by-surface basis to yield the digital and analog comparison estimates shown in @tbl-shop-comparisons. Overall, our algorithm would make `r format(sum(comparisons$Algorithm), big.mark=',')` (dependent) comparisons and an examiner doing a visual inspection would make `r format(sum(comparisons$Examiner), big.mark=',')` comparisons under the assumptions we made in Scenario 1.

@@

To fully understand the magnitude of this problem, suppose that we have a group of 4 additional individuals who may have contributed to building the un-detonated bomb. These individuals are each associated with the author and have hobbies that include building electronic devices, home improvement, welding, and woodworking (thus, they each have sets of tools). As part of the investigation, each individuals' tools are confiscated and examined; blade cuts are made of each cutting surface to be matched to the wire in evidence. @tbl-garage-comparisons shows the number of comparisons which would need to be performed to ensure that all wire-cutting tools in each suspect's garage are compared to the evidence from the scene.

```{r}
#| label: tbl-garage-comparisons
#| tbl-cap: "Total cutting length of easily accessible wire-cutting tools for 4 individuals associated with the author."
#| echo: false
#| message: false
set.seed(42309427)

n_tools <- c(40, 20, 30)
n_blades <- sample(5:75, size = 3)

tool_options <- select(bind_rows(cutting_surfaces, dad_surfaces), 1:3) %>% unique()

blades <- filter(tool_options, type %in% c("Utility knife blades", "Razor blades"))

other_tools <- filter(tool_options, !type %in% c("Utility knife blades", "Razor blades"))

dad_tools <- dad_surfaces %>% 
  mutate(suspect = 4) %>% select(-total_cutting_length)

our_tools <- cutting_surfaces %>% mutate(suspect = 0) %>% select(-total_cutting_length)

sim_garages <- tibble(suspect = 1:3, n_tools = n_tools, n_blades = n_blades) %>%
  mutate(data = map2(
    n_tools, n_blades, 
    ~bind_rows(
      slice_sample(other_tools, n = .x, replace = T), 
      slice_sample(blades, n = .y, replace = T)) 
  )) %>%
  unnest(data) %>%
  group_by(suspect, type, length, surfaces) %>% 
  summarize(tools = n()) %>%
  ungroup()

garages <- bind_rows(sim_garages, dad_tools, our_tools) %>%
  group_by(suspect) %>%
   mutate(digital = (length*10*mm_to_mum/resolution - wire_diam*mm_to_mum/resolution + 1)*surfaces*tools,
         analog = (length*10/examiner_increment - wire_diam/examiner_increment + 1)*surfaces*tools)

garage_sum <- garages %>%
  summarize(
    total_cutting_length = sum(length * surfaces * tools), 
    digital = sum(digital), 
    analog = sum(analog)) %>%
  mutate(suspect = as.character(suspect))

garage_sum_tot <- bind_rows(garage_sum, tibble(suspect = "Total", total_cutting_length = sum(garage_sum$total_cutting_length), digital = sum(garage_sum$digital), analog = sum(garage_sum$analog)))

garage_sum_tot %>%
  rename(Individual = suspect, `Total Cutting Length (cm)` = total_cutting_length, "Algorithm" = digital, "Examiner" = analog) %>%
  knitr::kable(format.args = list(big.mark = ','))
```

If investigators were to attempt to determine which of the suspected individuals cut the wire in evidence, they would need to do `r format(sum(garage_sum$analog), big.mark = ',')` distinct visual comparisons. Even if there is an incredibly small false positive error rate, the odds of a false identification are much more likely in this scenario than either of the previous two scenarios. But how likely? In the next section, we describe two different scenarios for calculating the overall likelihood of a false identification using simple probability calculations and estimated coincidental match probabilities as well as a theoretical approach designed to mimic the approach used with database searches that return the $N$ most similar results.

### 2023-11-17 Preparation/effective_obs.qmd
Updates

wce %>%
  ggplot(aes(x = x, y = sig)) + 
  geom_line() + 
  ggtitle("T1AW-LI-R1 A signature") + 
  theme_bw() + 
  ylab("Signature height") + xlab("Position")
```
No matter how small our resolution, the real-world surface is never more than 1.383 mm in length and the underlying striations are essentially fixed - what changes is what level of detail we can observe. 
At some point, adding additional resolution doesn't get us any more information about the striations; thus, we cannot depend solely on $n$, the number of observations, to tell us how much information is contained in our sequence.
In order to correct for this, time series analysts use an effective sample size calculation of $N^\prime = N \cdot \frac{1-r_1}{1+r_1}$ when there is first-order autocorrelation.

@@

wce_ts <- wce %>%
  mutate(sig1 = c(NA, diff(sig, 1)),
         sig2 = c(NA, diff(sig1, 1)))
model <- lm(sig ~ x, data = wce_ts)
lmtest::dwtest(model)
```
The null hypothesis of the Durbin-Watson test is that there is no autocorrelation; we can conclude that there is evidence of autocorrelation in our data.

>>>>

No matter how small our resolution, the real-world surface is never more than 1.383 mm in length and the underlying striations are essentially fixed - what changes is what level of detail we can observe. At some point, adding additional resolution doesn't get us any more information about the striations; thus, we cannot depend solely on $n$, the number of observations, to tell us how much information is contained in our sequence. In order to correct for this, time series analysts use an effective sample size calculation of $N^\prime = N \cdot \frac{1-r_1}{1+r_1}$ when there is first-order autocorrelation.

@@

wce_ts <- wce %>%
  mutate(sig1 = c(NA, diff(sig, 1)),
         sig2 = c(NA, diff(sig1, 1)))
model <- lm(sig ~ x, data = wce_ts)
lmtest::dwtest(model)
```

The null hypothesis of the Durbin-Watson test is that there is no autocorrelation; we can conclude that there is evidence of autocorrelation in our data.
